---
title: "10: Data Import"
output: github_document
---

This notebook will include all of the data importing steps, as well as cleaning and joining dataframes where necessary. I will need Tidyverse (primarily for dplyr and lubridate), janitor for name cleaning, and chron for getting uniform lap times. 

```{r load packages, message=F}
library(tidyverse)
library(janitor)
library(stringr)
```

First I will import the two qualifying datasets. They are split chronologically, with the first dataset including years 1950-2005 and the second 2006-2020. 

```{r import qualifying data, message=F, warning=F}
qual_06_20 <- read_csv("data/qualifying_times_2006-2020.csv") %>% clean_names()
```

```{r make qualifying datasets match, warning=F}
# Function to replace DNF and DNS with "NA"
clean_times <- function(x) ifelse(x %in% c("DNF", "DNS"), NA, x)

# Apply function to dataset
qual_06_20[9:11] <- lapply(qual_06_20[9:11], clean_times)

# Apply lubridate's 'ms' time conversion function to laptimes 
qual_06_20[9:11] <- lapply(qual_06_20[9:11], lubridate::ms)

# Finally, convert those columns to numeric
qual_06_20[9:11] <- sapply(qual_06_20[9:11], as.numeric)

# Check our results
view(qual_06_20)
```

Great, we have a nice clean dataframe ready for some exploration!