---
title: "20: Data Exploration"
output: github_document
---

First, let's just see how lap times compare across sessions. We would expect Q3 to have the quickest laps, followed by Q2 and then Q1. 

```{r view lap times by qualifying session}
density_plot <- function(x, y, z) {
  plot(density(na.omit(x)), 
       xlim=c(50,150),
       xlab = "Lap time (s)", main = "Lap times by qualifying session")
  
  par(new=T)
  plot(density(na.omit(y)), 
       xlim=c(50,150), 
       axes = F, xlab = "", ylab = "", main = "",
       col = "red", )
  
  par(new=T)
  plot(density(na.omit(z)), 
       xlim=c(50,150), 
       axes = F, xlab = "", ylab = "", main = "", 
       col = "blue")
  
  legend(130, 0.035, 
         legend = c("Q1", "Q2", "Q3"), 
         col = c("black", "red", "blue"), 
         lty=1)
}

density_plot(qual_06_20$q1, qual_06_20$q2, qual_06_20$q3)
```

Indeed, we see the trend we were expecting. Q3 features the quickest lap times of the three sessions.

Quicker drivers will go on to the subsequent sessions. Let's see how good of a predictor Q1 times are for eventual participation in Q3. 

<<<<<<< HEAD
```{r create column indicating q3 participation}
q3_participation <- function(x) ifelse(!is.na(x), 1, 0)
  
qual_06_20 <- qual_06_20 %>% 
  mutate(q3_y_n = as.factor(q3_participation(q3))) %>%
  filter(q1 < 1000)
=======
data_master %>% 
  filter(tyre %in% c("H", "M", "S")) %>% 
  ggplot(., aes(age,
                laptime_s)) +
  geom_point(aes(col = tyre)) +
  scale_color_manual(values=c("black", "gold", "red")) +
  geom_smooth(method = 'gam') +
  xlab("Tyre age (laps)") +
  ylab("Lap time (s)") +
  theme_classic()
>>>>>>> 9eb83abb46f5071c9d9166c9428f4fa2ce6322ca
```

```{r test relationship}
t1 <- qual_06_20$q1[qual_06_20$q3_y_n == 1]
t2 <- qual_06_20$q1[qual_06_20$q3_y_n == 0]

<<<<<<< HEAD
t.test(t1, t2)
=======
by_team <- function(driver_1, driver_2) {
  data_master %>% 
  filter(tyre %in% c("H", "M", "S"),
         driver %in% c(driver_1, driver_2)) %>% 
  ggplot(., aes(age,
                laptime_s)) +
  geom_point(aes(col = tyre)) +
  scale_color_manual(values=c("black", "gold", "red")) +
  geom_smooth(method = 'gam') +
  xlab("Tyre age (laps)") +
  ylab("Lap time (s)") +
  theme_classic()
}

rbr <- by_team("VER", "ALB")
merc <- by_team("HAM", "BOT")
rp <- by_team("PER", "STR")
mcl <- by_team("SAI", "NOR")
atau <- by_team("GAS", "KVY")
alfa <- by_team("GIO", "RAI")
haas <- by_team("GRO", "MAG")
ren <- by_team("RIC", "OCO")
fer <- by_team("VET", "LEC")
wil <- by_team("RUS", "LAT")

grid.arrange(rbr, merc, rp, mcl, atau, alfa, haas, ren, fer, wil, 
             nrow = 5)
>>>>>>> 9eb83abb46f5071c9d9166c9428f4fa2ce6322ca
```

These results suggest that Q1 times of those who eventually participate in the same GP's Q3 session are about 2 seconds quicker, on average, than those who do not end up qualifying for Q3. Let's visualize the distributions of these groups. 

<<<<<<< HEAD
```{r visualize relationship}
ggplot(data = qual_06_20, aes(x = q1, y = q3_y_n))+
  geom_jitter(height=0.2, alpha = 0.2)+
  geom_violin(draw_quantiles = 0.5, alpha = 0)+
  theme_classic()+
  xlab("Q1 lap times (s)")+
  ylab("Participation in Q3")
=======
```{r warning = F}
data_master %>% 
  filter(laptime_s > 100,
         tyre == "S") %>% 
  ggplot(., aes(x = age,
                y = laptime_s)) +
    geom_smooth(aes(col = team),
                se = F) +
    scale_color_brewer(palette="Paired") +
    theme_classic()
>>>>>>> 9eb83abb46f5071c9d9166c9428f4fa2ce6322ca
```

```{r}
redbull_sector_times <- data_master %>% 
  filter(s1 < 40,
         team == "Red Bull")

ggplot(data = redbull_sector_times, 
       aes(x = 1, 
           y = s1,
           col = driver)) +
  geom_boxplot() +
  geom_boxplot(data = redbull_sector_times, 
               aes(x = 2,
                   y = s2)) +
  geom_boxplot(data = redbull_sector_times, 
               aes(x = 3,
                   y = s3)) +
  scale_x_continuous(breaks = c(1, 2, 3)) + 
  xlab("Sector") +
  ylab("Sector time (s)") +
  ggtitle("Sector times, Red Bull Racing") +
  theme_classic()
```

```{r}
fun_sector_times <- data_master %>% 
  filter(s1 < 40,
         driver %in% c("ALB", "GAS"))

ggplot(data = fun_sector_times, 
       aes(x = 1, 
           y = s1,
           col = driver)) +
  geom_boxplot() +
  geom_boxplot(data = fun_sector_times, 
               aes(x = 2,
                   y = s2)) +
  geom_boxplot(data = fun_sector_times, 
               aes(x = 3,
                   y = s3)) +
  scale_x_continuous(breaks = c(1, 2, 3)) + 
  xlab("Sector") +
  ylab("Sector time (s)") +
  ggtitle("Sector times, Albon and Gasly") +
  theme_classic()
```

