---
title: "20: Data Exploration"
output: github_document
---

First, let's just see how lap times compare across sessions. We would expect Q3 to have the quickest laps, followed by Q2 and then Q1. 

```{r view lap times by qualifying session}
density_plot <- function(x, y, z) {
  plot(density(na.omit(x)), 
       xlim=c(50,150),
       xlab = "Lap time (s)", main = "Lap times by qualifying session")
  
  par(new=T)
  plot(density(na.omit(y)), 
       xlim=c(50,150), 
       axes = F, xlab = "", ylab = "", main = "",
       col = "red", )
  
  par(new=T)
  plot(density(na.omit(z)), 
       xlim=c(50,150), 
       axes = F, xlab = "", ylab = "", main = "", 
       col = "blue")
  
  legend(130, 0.035, 
         legend = c("Q1", "Q2", "Q3"), 
         col = c("black", "red", "blue"), 
         lty=1)
}

density_plot(qual_06_20$q1, qual_06_20$q2, qual_06_20$q3)
```

Indeed, we see the trend we were expecting. Q3 features the quickest lap times of the three sessions.

Quicker drivers will go on to the subsequent sessions. Let's see how good of a predictor Q1 times are for eventual participation in Q3. 

<<<<<<< HEAD
```{r create column indicating q3 participation}
q3_participation <- function(x) ifelse(!is.na(x), 1, 0)
  
qual_06_20 <- qual_06_20 %>% 
  mutate(q3_y_n = as.factor(q3_participation(q3))) %>%
  filter(q1 < 1000)
```

```{r test relationship}
t1 <- qual_06_20$q1[qual_06_20$q3_y_n == 1]
t2 <- qual_06_20$q1[qual_06_20$q3_y_n == 0]

t.test(t1, t2)
```

These results suggest that Q1 times of those who eventually participate in the same GP's Q3 session are about 2 seconds quicker, on average, than those who do not end up qualifying for Q3. Let's visualize the distributions of these groups. 


```{r visualize relationship}
ggplot(data = qual_06_20, aes(x = q1, y = q3_y_n))+
  geom_jitter(height=0.2, alpha = 0.2)+
  geom_violin(draw_quantiles = 0.5, alpha = 0)+
  theme_classic()+
  xlab("Q1 lap times (s)")+
  ylab("Participation in Q3")
```

